Mapping the characteristics of small-scale agricultural systems at national extents using small sat data and an active learning approach

### 2021-02-26 inst/manuscript/manuscript.Rmd
Full discussion drafted

@@

@@

- name: Meridia
  affilnum: '6'

@@

Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of arable croplands in Ghana. The cropland extent estimated from the vectorized map is 42,359 km$^2$ (with a margin of error of 4,395 km$^2$), or `r round(42359 / (42359 + 204940) * 100, 1)`% of the mapped area, the estimate based on the per pixel map is 43,233 km$^2$ (margin of error = 4,904 km$^2$), or `r round(43233 / (43233 + 202856) * 100, 1)`% of area.

@@

aoi_nofor <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
  summarize(weighted.mean(fsize, ntiles)) %>% pull() %>% round(., 2)
aoi_nofor_adj <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
  summarize(weighted.mean(fsizeadj, ntiles)) %>% pull() %>% round(., 2)
```

These results demonstrate our system's ability to create high resolution maps of the distribution and characteristics of cropland across a diverse and hard-to-map agricultural systems, at country-scales and at sub-annual to annual time steps. Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our 2018 maps, in which cropland covers 16.1-23.2% of the mapped area (Table \ref{tab:mapaccuracy}), and yield a sample-corrected estimate `r round(c(42359) / (42359 + 204940) * 100, 1)` to `r round(c(43233) / (43233 + 202856) * 100, 1)`%. This result could suggest that Ghana's arable cropland occupies is less than these previous estimates, or that these previous maps used a more expansive definition of cropland than we used.

@@

## Sources of error
Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors. The overall map accuracies are good (86.7-88%, Table \ref{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence). However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map, indicating that both commission and omission errors were above 30% for this class. The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%). These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017]

These relatively low accuracies for the cropland class can be attributed to several important factors. The first of these was the overall mapping extent combined with the resolution of the analysis. Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended [@MaxwellImplementationmachinelearningclassification2018] tree number of depth  in our RandomForest implementation. Previous work shows that RandomForest achieves much better performance on small-scale croplands when trained on a large number of randomly generated features [@Debatsgeneralizedcomputervision2016], which is a commonality with deep learning. However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the typical 10-14 hours required from training through prediction on 800 CPUs, for a single iteration. This left the model less able to discern cropland from surrounding savanna vegetation with bare patches or sparse grass cover, which was common in most AOIs. A larger feature set with more variable textures filters could have helped distinguish such areas from crop fields, helping to particularly reduce commission error.

@@

## Key system features
Despite the sources of errors mentioned above, the design of the system proved effective in mitigating error, resulting in a higher overall accuracy than would have otherwise been possible. The label accuracy assessment and consensus labelling methods appeared to be the most effective tools for error mitigation. Label accuracy assessment allowed us to quantify the impact of trainnig data error on model performance, in keeping with recommended best practices [@ElmesAccountingtrainingdata2020]. Label quality proved to have a large impact on model performance (Figure \ref{fig:trainingimpact}), and consensus labeling was effective in reducing individual digitization errors and between-labeller differences in interpretation, improving label, and thus map, quality. Worker-specific label accuracy measures also helped to improve the overall quality of the consensus labels, by enabling higher weight to be placed on labels that were more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels. The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we could in turn use to bias-correct map-based size class estimates (Table \ref{tab:sizentab}). The ability to select the most accurate label will also help when different training strategies are needed, such as tasking a model to classify both field edge and interior classes [e.g. @waldnerDeepLearningEdge2020; @aungFarmlandParcelDelineation2020]. In such cases, the vectorized labels are needed to preserve information about the field edges, which are removed during consensus label generation.

@@

Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpacts}) was influenced by the validation labels we used, which were generated using the consensus method. This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels. However, the visual assessment of probability images generated from each labelling strategy show that these two strategies produce nearly identical maps, and that the model trained with low quality labels generated a markedly less certain map (Figure S6).

## Looking forward 
- Provide new insight 
- Given errors, should be considered generation 1 maps
- Sets up capability for repeated annual mapping of croplands to understand critical issues of how these systems are changing (sizes and shapes of croplands, where they are expanding) (Jayne papers, recent agricultural expansion paper)
- Key components will be useful in other applications and regions
- Provides means for improving crop type and crop yield estimates
- Improvements to help realize these capabilities: 
-     [This suggests that replacing the Random Forest more advanced deep learning model]. promise of CNNs (cite Waldner paper, others), for segmentation to realize this. This systems is well designed to adapt to it, and capabilities for assessing various aspects of label quality, e.g. edge accuracy and fragmentation accuracy, can help improve quality of labels designed to emphasize edges
  - Can still retain the active architecture, but instead use new sites to improve a pre-trained model
  - Deep learning may be more resistant to label errors (cites)
- Map along agroecozone boundaries, rather than random blocks  
- Another interesting question to answer: whether guided approach to site selection would result in more informative sample for CNNs also.
- Higher number of high quality reference labels can achieve this.

- Next steps: 
    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.

# Acknowledgements

>>>>

- name: Zhenhua Meng
  affilnum: '1'

- name: Qi Zhang
  affilnum: '1'

@@

- name: Justus Muhando
  affilnum: '5'

@@

@@

```{r, echo=FALSE}
data("accuracy_reports")

# Adjusted areas
rfa <- accuracy_reports$ghana$rf[, "Area"]
rfa <- rfrep[nchar(rfa) > 0]
rfas <- as.numeric(gsub("\\(.*", "", rfa[nchar(rfa) > 0]))
rfma <- as.numeric(stringr::str_match(rfa[1], "\\(\\s*(.*?)\\s*\\)")[, 2])
cipct <- function(x, moe) {
  apct <- x[2] / sum(x)
  moepct <- moe / sum(x)
  round(c(apct, apct - moepct, apct + moepct) * 100, 1)
}
sega <- accuracy_reports$ghana$seg[, "Area"]
sega <- segrep[nchar(sega) > 0]
segas <- as.numeric(gsub("\\(.*", "", sega[nchar(sega) > 0]))
segma <- as.numeric(stringr::str_match(sega[1], "\\(\\s*(.*?)\\s*\\)")[, 2])

rfcipct <- cipct(rfas, rfma)
segcipct <- cipct(segas, segma)

fmt <- function(x) format(x, big.mark = ",")
```

Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of arable croplands in Ghana. The cropland extent estimated from the vectorized map is `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area. The estimate based on the per pixel map is `r fmt(rfas[2])` km$^2$ (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.

@@

aoi_nofor <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
  summarize(weighted.mean(fsize, ntiles)) %>% pull() %>% round(., 2)
aoi_nofor_adj <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
  summarize(weighted.mean(fsizeadj, ntiles)) %>% pull() %>% round(., 2)

These results demonstrate our system's ability to create high resolution maps of the distribution and characteristics of cropland across a diverse and hard-to-map agricultural systems, at country-scales and at sub-annual to annual time steps. Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our 2018 maps, in which cropland covers 16.1-23.2% of the mapped area (Table \ref{tab:mapaccuracy}), with sample-corrected estimates of `r segcipct[1]`-`r rfcipct[1]`%. Our findings suggests that Ghana's arable cropland is slightly less than previous estimates, although the difference may also be attributable to the earlier maps using a more expansive definition of cropland than ours (active to recently active fields).

@@

## Sources of error
Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors. The overall map accuracies are good (86.7-88%, Table \ref{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence). However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map across the country, indicating that both commission and omission errors were above 30% for this class. The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%). These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

The patterns of cropland-class accuracies varied by zone, which largely align with agroecozones and thus somewhat reflective of the inherent challenge each poses for mapping. Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savanna agroecoregions (Figure S9), and lowest in zones 3 and 4, which are either forest or coastal savannas. User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample. Confining the samples to specific agroecozones provides further insight into error patterns (Table S4). The coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably due to the numerous areas of high density cropland and low woody cover in surrounding uncultivated areas, which promotes commission error. The northern savannas had the best balance between omission and commission error, and had the highest overall User's accuracy. The transition zone had a very low Producer's accuracy (21%), which likely reflects the fact that this region, which also has high cropland density, was divided between several AOIs for mapping.

The relatively low accuracies for the cropland class can be attributed to several important factors. The first of these was the overall mapping extent combined with the resolution of the analysis. Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forest implementation. Previous work shows that RandomForest achieves much better performance on small-scale croplands when trained on a large number of randomly generated features [@Debatsgeneralizedcomputervision2016], which is a commonality with deep learning. However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the typical 10-14 hours required from training through prediction on 800 CPUs, for a single iteration. This left the model less able to discern cropland from surrounding savanna vegetation with bare patches or sparse grass cover, which was common in most AOIs. A larger feature set with more variable textures filters could have helped distinguish such areas from crop fields, helping to particularly reduce commission error.

@@

## Key system features
Despite the sources of errors mentioned above, the design of the system proved effective in mitigating error, resulting in a higher overall accuracy than would have otherwise been possible. The label accuracy assessment and consensus labelling methods appeared to be the most effective tools for error mitigation. Label accuracy assessment allowed us to quantify the impact of trainnig data error on model performance, in keeping with recommended best practices [@ElmesAccountingtrainingdata2020]. Label quality proved to have a large impact on model performance (Figure \ref{fig:trainingimpact}), and consensus labeling was effective in reducing individual digitization errors and between-labeller differences in interpretation, improving label, and thus map, quality. Worker-specific label accuracy measures also helped to improve the overall quality of the consensus labels, by enabling higher weight to be placed on labels that were more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels. The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we could in turn use to bias-correct map-based size class estimates (Table \ref{tab:sizentab}). The ability to select the most accurate label will also help when different training strategies are needed, such as when training a model to perform instance segmentation [@carvalhoInstanceSegmentationLarge2021], in which case the field identities have to be preserved.

@@

Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpact}) was influenced by the validation labels we used, which were generated using the consensus method. This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels. However, the visual assessment of probability images generated from each labelling strategy show that these two strategies produce nearly identical maps, and that the model trained with low quality labels generated a markedly less certain map (Figure S6).

## Next steps
The maps presented here represent a version 1 product that is freely available to use **(links)**, along with its underlying code. These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020]. In their current form, the maps may be useful for a variety of research applications. For example, analyzing the  distributions of values in the probability maps may provide additional insight into the relative extents of active versus fallow croplands [@Tongforgottenlanduse2020]. However, use of these data, particularly for decision-making processes (e.g. cropped area estimates), should be careful to account for the reported errors [@OlofssonGoodpracticesestimating2014; @StehmanKeyissuesrigorous2019].

To facilitate the next step, generating more accurate version 2 maps, several improvements will be made. The first is to replace Random Forest with a more advanced convolutional neural network (CNN), which can generate and learn from a large number of features representing a variety of spatial scales [@maDeepLearningRemote2019].  Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, provide an effective means for delineating field boundaries [@waldnerDeepLearningEdge2020]. Our system can readily incorporate such a model. The active learning framework is already analagous to transfer learning, in which a trained deep learning model is applied to  a different region and updated with local training samples [@panSurveyTransferLearning2010; @russwurmMetalearningFewshotLand2020]. The existing tooling can thus be adjusted to guide training site selection in sub-regions of Ghana for a CNN trained from scratch from a larger initial random draw. Instead of the current AOIs, a more effective strategy may be to update the model within agroecozones (Figure S9B) The labelling methods already provide for a case in which edge and interior classes need to be labelled, as the most accurate individual training samples can be identified and used, rather than the consensus strategy. The segmentation algorithm will also be updated to work with the resulting score maps.

## Conclusion
This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands over national to regional scales, using a flexible and framework that can be readily updated to improve their accuracy as technologies improve. Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.

# Acknowledgements
The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund. Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393). We thank Joe Okyere and his colleagues at Meridia for assistance in data collection.
